### 2019-04-18 code/rmarkdown/Supporting_information.rmd
first commit

---
title: "Supporting Information: Phylogeographic inference with microsatellites and SNPs"
author: |
  | Miguel Camacho-Sanchez${\ast^1}$, Guillermo Velo-Antón$^{1}$, Ana Veríssimo$^1$, Silvia Carvalho$^2$
  | $^1$ CIBIO/InBIO, Centro de Investigação em Biodiversidade e Recursos Genéticos da Universidade do Porto, Vairão, Portugal
  | Correspondence should be addressed to miguelcamachosachez@gmail.com
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmarkdown::pdf_document:
    toc: false
    keep_tex: true
    fig_caption: yes
    includes:
      in_header: preamble.tex
fontsize: 11pt
documentclass: article
subparagraph: true
---