### 2019-04-18 code/rmarkdown/article.Rmd

---
title: "Phylogeographic inference with microtellite and SNPs"
author: |
  | Miguel Camacho-Sanchez${\ast^1}$, Guillermo Velo-Antón$^{1}$, Ana Veríssimo$^1$, Silvia Carvalho$^2$
  | $^1$ CIBIO/InBIO, Centro de Investigação em Biodiversidade e Recursos Genéticos da Universidade do Porto, Vairão, Portugal
  | Correspondence should be addressed to miguelcamachosachez@gmail.com
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmarkdown::pdf_document:
    toc: false
    keep_tex: true
    fig_caption: yes
    includes:
      in_header: preamble.tex
fontsize: 11pt
documentclass: article
subparagraph: true
---

```{r, include = FALSE}
knitr::opts_knit$set(root.dir = normalizePath("../.."))
```

```{r, include = FALSE}
# restore session
gen <- readRDS("data/intermediate/raw_genotypes.RDS")
# load packages
# define helper functions
# load data

```

# MANUSCRIPT INFORMATION

\clearpage

# ABSTRACT

TODO

\clearpage

# 1 | INTRODUCTION

# 2 | MATERIALS AND METHODS

## Study area and species

## Data

# 3 | RESULTS

* result 1: sequencing
We sequenced a total of ´r adegenet::nInd(gen)´ samples which included 1 replica. DArTseq genotyped ´r adegenet::nLoc(gen)´SNPs in ´r nlevels(as.factor(gen@other$loc.metrics$CloneID))´ loci. The overall missing data was ´r ´ ()

* result 2: structure vs dapc

* result 3: genetic diversity

# 4 | DISCUSSION

* Overview

* Key result 1

* Key result 2

* Key result 3

* Concluding paragraph

# AUTHORS' CONTRIBUTIONS

TODO

# DATA ACCESSIBILITY

# ORCID

Miguel Camacho-Sanchez: 0000-0002-6385-7963 \newline
Guillermo Velo-Antón: 0000-0002-9483-5695 \newline
Ana Veríssimo: 0000-0003-3396-9822 \newline
JEFF ADAM PEDRO?
Silvia B. Carvalho: 0000-0003-4368-4708 \newline

# REFERENCES