### 2019-06-26 code/rmarkdown/Figures.rmd
intro and Fig1

---
title: "Figures"
fontsize: 11pt
documentclass: article
subparagraph: true
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: no
    includes:
      in_header: preamble.tex
---

```{r, include = FALSE}
knitr::opts_knit$set(root.dir = normalizePath("../.."))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = "png", dev.args = list(type = "cairo"), dpi = 500, fig.align = "center")
```

##Figure 1:
```{r, fig.height = 4.5, fig.width = 4.5, fig.cap = "Populations sampled in *Hyla molleri* (panel A) and *Pelobates cultripes* (panel B), for microsatellite (triangles) and SNP (circles). Coordinates and samples codes are in Table 1. The distribution for each species is indicated in shaded area, downloaded from the IUCN on May 2019 (*H. molleri* was described recently and its distribution is not availabe at the IUCN database. We used the distribution of *H. arborea* instead)"}

```