### 2019-06-28 code/rmarkdown/tables.Rmd
table 1

---
title: "Tables"
output:
  word_document: default
  html_document:
    df_print: paged
---
Table 1. Samples and their localities. In the sample column, we indicate the species(*P.c.* and *H.m*, to refer to *Pelobates cultripes* and *Hyla molleri*, respectively) and marker type (SNPs or microsatellite).
```{r, eval = T, echo = F, include = T, warning = F, message=FALSE}
library(flextable)
library(dplyr)
h <- readRDS("../../data/intermediate/table1.rds")

ft <-
h %>% flextable::flextable() %>%
  flextable::width(j = 5, 3.7) %>%
  flextable::width(j = 2, 2) %>%
  flextable::fontsize(size = 8, part = "body") %>%
  flextable::fontsize(size = 10, part = "header")
ft
```