### 2019-07-01 code/rmarkdown/Supporting_information.rmd
convergence plots

---
title: "Supporting Information: Phylogeographic inference with microsatellites and SNPs"
author: |
  | Miguel Camacho-Sanchez${\ast^1}$, Guillermo Velo-Antón$^{1}$, Ana Veríssimo$^1$, Silvia Carvalho$^2$
  | $^1$ CIBIO/InBIO, Centro de Investigação em Biodiversidade e Recursos Genéticos da Universidade do Porto, Vairão, Portugal
  | Correspondence should be addressed to miguelcamachosachez@gmail.com
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmarkdown::pdf_document:
    toc: false
    keep_tex: true
    fig_caption: yes
    includes:
      in_header: preamble.tex
fontsize: 11pt
documentclass: article
subparagraph: true
---
Table S1: samples
Figure S1: SNP filtering for *H. molleri* (A, C, E, G) and *P. cultripes* (B, D, F, H). En each panel we show the frequency distribution (y-axes) of the metric evaluated (x-axes), with the thresholds applied (dashed vertical lines).

Figures 2.1 - 2.11: Change of the likelihood of the data across iterations for 10 replicate runs for K = 1 to K = 10. The vertical dashed lines indicate burnin length.

Figures S3.1 - S3.2: Gelman and Rubin's convergence diagnostic for SNP (S3.1) and microtellite (S3.2) datasets.